---
title: "Statistical Inference Report"
author: "Anandu R"
date: "9/20/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=5, fig.height=4)
```

Within this exercise using statistical inference and knowledge on various 
distributions, I investigate the exponential distribution and compare it with 
the Central Limit Theorem. I will attempt to relate the two distributions -
Exponential distribution and Uniform distribution.

Creating an exponential distribution
```{r}
set.seed(13)
lambda = 0.2
n = 40 
simulations = 1000
dstbn = rexp(n, lambda)
```

We will now use bootstrap sampling to create a sample distribution
```{r}
dstbn_sampled = matrix(sample(dstbn,simulations*n,replace = T),simulations,n)
dstbn_resampled = apply(dstbn_sampled, 1, mean)
```

## Comparing the means
```{r, echo=F, results=F}
theory_mean = 1/lambda
observed_mean = mean(dstbn_resampled)
``` 
```{r, echo = F}
data.frame(distribution = c("Theoretical","Observed"),means = c(theory_mean,observed_mean))
```  

Visualization
```{r}
hist(dstbn_resampled, xlab = "mean", main = "Exponential Function Simulations", col = "hotpink3")
abline(v = theory_mean, col = "skyblue", lwd = 3)
abline(v = observed_mean, col = "lightyellow3", lwd = 3)
legend("topright", col = c("skyblue","lightyellow3"), legend = c("Theoretic mean", "Observed mean"), pch = 19)
```


The theoretical mean is `r round(theory_mean,2)` whereas the observed mean of the 40 averages from
the 1000 sampled distributions is `r round(observed_mean,2)`

## Comparing variability 
```{r, echo=F, results=F}
theory_var = (theory_mean/sqrt(n))^2
observed_var = var(dstbn_resampled)
``` 
Comparing variability in the sampled distribution compared to the theoretical
distribution variance.
```{r, echo=F}
data.frame(distribution = c("Theoretical","Observed"),variance = c(theory_var,observed_var))
```  
The theoretical variance is `r round(theory_var,4)` whereas the observed variance of the 40 averages from
the 1000 sampled distributions is `r round(observed_var,4)`

## Visualising the sampled distribution
Representation of the sample distribution as a density plot to investigate the
normality of the distribution
```{r}
## Creating dataframe for visualization
df = data.frame(obs=c(rep("observed",40),rep("sampled",40)),value=c(dstbn,dstbn_resampled))

## Visualization
library(ggplot2)
ggplot(
  df,
  aes(
    x = value,
  )
) + geom_density(fill = "hotpink3", color = "red") +
  theme_bw() +
  coord_cartesian(xlim = c(0,10)) + 
  ggtitle("Density of sample distribution")
  
```   

We can observe that the distribution has a bell-shaped curve, a classical 
feature of normal distribution.

### Approximating the sample distribution to a standard normal
Normalizing the sample means using z-scores  
<img src="https://render.githubusercontent.com/render/math?math=$\mathrm{z} = \frac{x-1/\lambda}{\frac{1/\lambda}{\sqrt{n}}}$">  

According to the central limit theorem the z-scores should be centered around the 
mean of the standard normal, 0.

```{r}
dstbn_normalized = data.frame(z=as.numeric((dstbn_resampled-theory_mean)/sqrt(theory_var)))
z_mean = mean(dstbn_normalized$z)
```  

Visualizing
```{r}
ggplot(
  dstbn_normalized,
  aes(
    x = z
  )
) + geom_histogram(alpha=0.5,  binwidth = 0.3, fill="hotpink3", color="black", aes(y = ..density..)) +
   stat_function(fun = dnorm, size = 1.2) +
   geom_vline(xintercept = z_mean, color="red", size = 1.2) +
   ggtitle("Distribution of standardized simulated means") +
   xlab("z-scores")
```  
We see that the distribution falls almost within the standard normal 
distribution, and hence approximated.


