---
title: "PM2.5 Emissions Data Analysis"
author: "Anandu R"
date: "8/1/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PM2.5 Emissions Data Analysis
Fine particulate matter (PM2.5) is an ambient air pollutant for which there is 
strong evidence that it is harmful to human health. In the United States, the
Environmental Protection Agency (EPA) is tasked with setting national ambient 
air quality standards for fine PM and for tracking the emissions of this
pollutant into the atmosphere. Approximatly every 3 years, the EPA releases its
database on emissions of PM2.5. This database is known as the National Emissions
Inventory (NEI). You can read more information about the NEI at the EPA National
Emissions Inventory web site.

For each year and for each type of PM source, the NEI records how many tons of
PM2.5 were emitted from that source over the course of the entire year. The data
that is used in the analysis are for 1999, 2002, 2005, and 2008.

## Loading the data
#### Getting the data 
```{r}
fileUrl = "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2FNEI_data.zip"
if(!file.exists(
  "./data/summarySCC_PM25.rds"
  )|!file.exists(
    "./data/Source_Classification_Code.rds"
    )
  ){
  download.file(fileUrl,"data.zip")
  unzip("data.zip",exdir = "./data")
  unlink("data.zip")
}
```

#### Reading the data
```{r}
NEI <- readRDS("./data/summarySCC_PM25.rds")
SCC <- readRDS("./data/Source_Classification_Code.rds")
```

#### Preliminary analysis of data

NEI
```{r}
head(NEI)[,c(1,2,4:6)]
```
SCC 
```{r}
head(SCC)[,c(1,3)]
```



## Exploratory Analysis 

#### Visualization of total emissions per year from all sources
Using the base plotting system, to make a plot showing the total PM2.5 emission 
from all sources for each of the years 1999, 2002, 2005, and 2008.
```{r}
dat_total = tapply(NEI$Emissions,factor(NEI$year),sum)
dates = c(1999,2002,2005,2008)
```

Visualising the data
```{r}
plot(dates,
     dat_total,
     cex=1.5,
     col = 'red',
     pch =19,
     xlab = "Years",
     ylab = "Emissions",
     main = "Visualization of total emissions per year from all sources")
lines(dates,as.numeric(dat_total),lwd=2, col = "orange")
```
As we can see from the visualization above the emission levels have dropped from 
1999 to 2008

#### Examining the trend of emissions over time in Baltimore City
Using the base plotting system, to make a plot showing the total PM2.5 emission 
in Baltimore City alone for each of the years 1999, 2002, 2005, and 2008.

Subsetting the dataset 
```{r}
dat_balt = subset(NEI, fips == "24510")
em_balt = as.numeric(tapply(dat_balt$Emissions,factor(dat_balt$year),sum))
dat_balt_df = data.frame(dates, Emissions = em_balt)
```

Visualising the data
```{r}
plot(
  dat_balt_df,
  cex=1.5,
  col = 'red',
  pch =19, 
  xlab = "Years",
  ylab = "Emissions",
  main = "Examining the trend of emissions over time in Baltimore City")
lines(dat_balt_df,lwd=2, col = "orange")
```
We observe that within Baltimore City, the emission levels decrease from 1999 to 
2002 then the graph shows an increase from 2002 to 2005 but the emission levels
decrease considerably from 2005 to 2008

#### Indicating trends in emissions of different Source Types in Baltimore City 
There are various source types - point, non-point, onroad, non-road, we want to 
see which of these source types have had their emissions lowered over the years 
and which of them had increase in their emission levels.

Recasting the data based on need
```{r}
library(reshape2)
data_melt = reshape2::melt(dat_balt[,4:6],id.vars = c('year','type'))
data_cast = reshape2::dcast(data_melt, year+type~ variable, fun.aggregate = sum)
```

Visualising the data
```{r}
library(ggplot2)
ggplot(
  data_cast,
  aes(
    x = factor(year),
    y = Emissions,
    color = type,
    group = type
  ),
) + geom_point(cex = 2) +
  geom_line(cex =1.4) + 
  theme_bw() + guides(fill=FALSE)+
  labs(x="Year", y=expression("Total PM"[2.5]*" Emission (Tons)")) + 
  labs(title=expression("PM"[2.5]*" Emissions, Baltimore City 1999-2008 by Source Type"))
```
From the graph above we observe that the emission levels go down overtime for
the non-point, non-road and on-road souces, while that of point source has 
increased from 1999 to 2008

#### Emissions from coal combustion-related sources
We are to interpret how emissions from coal combustion-related sources have changed 
from 1999–2008 across the United States.

We identify the SCC value(s) that represent the coal combustion related sources

We need to identify souces that employ combustion of coal, the process - 
combustion, is identified in the SCC.Level.One whereas the fuel - coal, is 
identified by looking at the SCC.Level.Four variable. For the purpose we 
identify the SCC codes that represent the required data from the SCC dataset.

```{r}
combustionRelated = grepl("comb", SCC[, 'SCC.Level.One'], ignore.case=TRUE)
coalRelated = grepl("coal", SCC[, 'SCC.Level.Four'], ignore.case=TRUE) 
both = combustionRelated&coalRelated
reqSCC = SCC[both,'SCC']
```

Subsetting the dataset 
```{r}
dat_total_cc = subset(NEI, SCC %in% reqSCC)
dat_total_cc = tapply(dat_total_cc$Emissions,factor(dat_total_cc$year),sum)
```

Visualising the data
```{r}
plot(dates,
     dat_total_cc,
     cex=1.5,
     col = 'red',
     pch =19,
     xlab = "Years",
     ylab = "Emissions",
     main = "Emissions from coal combustion-related sources")
lines(dates,as.numeric(dat_total_cc),lwd=2, col = "orange")
```
We observe that the overall emissions do decrease from 1999 to 2008 eventhough
there is slight increase from 2002 to 2005.

#### Emissions from motor vehicle sources Baltimore City
To interpret how emissions from motor vehicle sources have changed from 
1999–2008 in Baltimore City, First we must identify the SCC codes that represent 
motor vehicle sources from the SCC dataset.

```{r}
motoVehicleRelated = grepl("vehicle", SCC[, 'SCC.Level.Two'], ignore.case=TRUE)
reqSCC = SCC[motoVehicleRelated,'SCC']
```

Subsetting the dataset
```{r}
dat_balt_mv = subset(dat_balt, SCC %in% reqSCC)
dat_balt_mv = tapply(dat_balt_mv$Emissions,factor(dat_balt_mv$year),sum)
```

Visualising the data
```{r}
plot(dates,
     dat_balt_mv, 
     cex=1.5,
     col = 'red',
     pch =19,
     xlab = "Years",
     ylab = "Emissions",
     main = "Emissions from motor vehicle sources Baltimore City")
lines(dates,as.numeric(dat_balt_mv), lwd = 2, col = "orange")
```
#### Comparing the emissions between Baltimore City and Los Angeles County
Compare emissions from motor vehicle sources in Baltimore City with emissions 
from motor vehicle sources in Los Angeles County, California (fips == "06037"). 
To interpret Which city has seen greater changes over time in motor vehicle
emissions.

Creating new dataset with the required information
```{r}
dat_losAngeles = subset(NEI, fips == "06037")
em_losAngeles = as.numeric(tapply(dat_losAngeles$Emissions,factor(dat_losAngeles$year),sum))
dat_balt_losAngeles_df = data.frame(dates, Emissions.Baltimore = em_balt, Emissions.LosAngeles = em_losAngeles)
head(dat_balt_losAngeles_df)
rng = range(dat_balt_losAngeles_df$Emissions.Baltimore,dat_balt_losAngeles_df$Emissions.LosAngeles)
```

Visualising the data 
```{r}
plot(
  ylim = rng,
  x = dates,
  type = "n",
  xlim = c(1998,2009),
  xlab = "Years",
  ylab = "Emissions",
  main = "Compare Baltimore City and Los Angeles County emissions")
points(dates,dat_balt_losAngeles_df$Emissions.Baltimore, cex = 1.5, col = "red", pch=19)
points(dates,dat_balt_losAngeles_df$Emissions.LosAngeles, cex = 1.5, col = "purple", pch=19)
lines(dates,dat_balt_losAngeles_df$Emissions.Baltimore, lwd = 2, col = "orange")
lines(dates,dat_balt_losAngeles_df$Emissions.LosAngeles, lwd = 2, col = "blue")
legend("topright", legend=c("Baltimore City","Los Angeles"),col = c("red","purple"),pch =19)
```
Comparing the data between that of Baltimore City and Los Angeles we observe 
that the emission levels are relatively the same for Baltimore City whereas the
emissions of Los Angeles has varied the most from 1999 to 2008 and has 
considerably decreased as well.

#### Removing files after analyis 
```{r}
unlink("data", recursive = T)
```



