---
title: "Colors in R plots"
output: html_document
---

importing the grDevices package components - colorRamp and colorRampPalette
```{r}
if(!require(grDevices)){
  install.packages("grDevices")
}
library(grDevices)
```

colorRamp: Take a palette of colors and return a function that takes values between 0 and 1, indicating the extremes of the color palette (e.g. see the 'gray function) 
colorRampPalette: Take a palette of colors and return a function that takes integer arguments and returns a vector of colors interpolating the palette (like heat.colors or topo.colors)

Creating a palette using blue and red to understand the colorRamp function
```{r}
pal = colorRamp(c("blue","red"))
```

The function pal now can be used to create colors using the two blotches blue and red
pal(1) returns the color blue
```{r}
pal(1)
```
Whereas pal(0) returns color red
```{r}
pal(0)
```
and similarly pal(0.5) returns purple, the color inbetween red and blue, or the color obtained by mixing the blotches red and blue
```{r}
pal(0.5)
```
Thus we can obtain various colors, with varying degree of the color blotches in the palette specified 
```{r}
pal(seq(0,1,length = 10))
```

The colorRampPalette works in similar way, but it returns a function that can be used to generate n number of colors using the color blotches specified 
```{r}
pal = colorRampPalette(c("blue","red"))
```

colorRampPalette(2) returns the two colors specified in the color blotch list
```{r}
pal(2)
```

and so on, passing any arbitray number to the function generates n colors that can be obtained by proportionately varying the intensity of the colors listed 
```{r}
pal(10)
```
pal(3) returns 3 colors, where the middle color represents purple
```{r}
pal(3)
```

The RColorBrewer package on CRAN contains interesting and useful color palettes 
```{r}
if(!require(RColorBrewer)){
  install.packages("RColorBrewer")
}
library(RColorBrewer)
```

There are three types of palettes available:
- Sequential : Used for data that are ordered, numeric or continuous data
- Diverging : Used for data that deviate or diverge 
- Qualitative : Used for data that are not ordered, they can be factors or discrete data 

The palette information can be used in conjunction with the colorRamp() and colorRampPalette() functions 

Usage: 

brewer.pal() function is used to return the colors we need for our purpose
it takes two arguments: an integer value which specifes the number of colors we want the function to return 
and, a string to specify the palette from which we want to extract the colors 
The names of various palettes can be obtained from the help page of brewer.pal() function

```{r}
cols = brewer.pal(3, "BuGn")
cols
```
And then we use these colors to generate our custom color using the colorRamp() or colorRampPalette() functions
```{r}
pal = colorRamp(cols)
pal(0.4)
```
```{r}
pal = colorRampPalette(cols)
pal(6)
```

Using the pal() to plot the volcano dataset 
```{r}
image(volcano, col = pal(20))
```

The smoothscatter function is used to plot a scatterplot of a lot of datapoints, the points are blended together to get an abstract image representation of the data, we can use custom palettes here to make the plots look better

general scatter plot: 
```{r}
set.seed(10)
x = rnorm(1000)
y = rnorm(1000)
plot(x,y)
```
smoothscatter(): 
```{r}
smoothScatter(x,y)
```


Colors can also be generated using the rgb function, by specifying the proportions of red, blue and green.
The transparency can be altered by specifying a value for the "alpha" parameter 

```{r}
plot(x,y,col = rgb(0,0,0,0.2),pch = 19)
```

