---
title: "Quantified Self Movement Monitor Data Analysis"
author: "Anandu R"
date: "8/3/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

It is now possible to collect a large amount of data about personal movement
using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up.
These type of devices are part of the “quantified self” movement – a group of
enthusiasts who take measurements about themselves regularly to improve their
health, to find patterns in their behavior, or because they are tech geeks. But 
these data remain under-utilized both because the raw data are hard to obtain
and there is a lack of statistical methods and software for processing and 
interpreting the data.

#### Loading the data 

Downloading the data
```{r}
fileUrl = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if(!file.exists("./data/activity.csv")){
  download.file(fileUrl, destfile = "./data/data.zip")
  unzip("./data/data.zip", exdir = "./data")
}
unlink("./data/data.zip", recursive = T)
```

Reading the raw data
```{r}
data = read.csv("./data/activity.csv")
```

Preliminary analysis on data
```{r}
dim(data)
```
```{r, echo=F}
head(data)
```
There seems to be a bit of missing values in the data.  

Taking a look at how much of the data is missing
```{r, results=F}
temp = mean(is.na(data))
```
which is just `r temp` of the data, we might as well get rid of the specific records 
which have missing values. 
```{r, echo=F,results=F}
good_data = complete.cases(data)
temp = mean(good_data)
```
calculating the % of records that have valid activity measurement, we find that 
`r temp` of data is good, so we get rid of rest of the records that have 
missing values
```{r, echo=F}
table(good_data)
```
```{r}
data_clean = data[good_data,]
## Corrected datatype of date field
data_clean = transform(data_clean, date= as.Date(date))
```


Taking a look at the valid activity tracked, for all 'steps'>0  
```{r, echo=F}
head(data_clean[data_clean$steps!=0,])
```

### Histogram of the total number of steps taken each day
```{r}
library(dplyr)
data_grpd = data_clean %>% 
  group_by(date) %>%
  summarise(steps_per_day = sum(steps))
data_grpd
```

```{r}
library(ggplot2)
ggplot(
  data_grpd,
  aes(
    x = steps_per_day,
  ),
) + 
  geom_histogram(bins = 12, fill = "skyblue", col = "white") +
  stat_bin(bins = 12,geom="text", aes(label=..count..), col ="white", vjust = 1.3) +
  theme_bw() + guides(fill = F) + 
  labs(
    title = "Histogram of the total number of steps taken each day",
    x = "Steps taken each day",
    y = "Frequency")
```
On average people take 10000 steps daily.

### Mean and median number of steps taken each day
```{r}
data_mm = suppressMessages(
  {
    data_grpd %>% 
      group_by(month = format(date, format= "%m")) %>%
      summarise(
        mean_steps = mean(steps_per_day),
        median_steps = median(steps_per_day)
      )
    }
  )
data_mm = transform(data_mm, month = month.abb[as.numeric(month)])
data_mm
```

### Time series plot of the average number of steps taken
```{r}
ggplot(
  data_grpd,
  aes(
    x = date,
    y = steps_per_day,
  )
) + 
  geom_line(col = 'skyblue', lwd = 1.25) + 
  theme_bw() + guides(fill = F) +
  labs(
    title = "Time series plot of the average number of steps taken",
    x = "Dates",
    y = "Step count"
  )

```
### Step count average in day across the time period 
Making a summarised data with step count average in day across the time period 
```{r}
data_consdt = data_clean %>% 
  group_by(interval) %>%
  summarise(steps_by_interval = mean(steps))
data_consdt
```

```{r}
idx_max_meanSteps = which.max(data_consdt$steps_by_interval)
max_meanSteps_interval = data_consdt$interval[idx_max_meanSteps]
ggplot(
  data_consdt,
  aes(
    x = interval,
    y = steps_by_interval,
  )
) + 
  geom_line(col = 'skyblue', lwd = 1.25) + 
  theme_bw() + guides(fill = F) +
  labs(
    title = "Time series plot of the average number of steps taken",
    x = "Dates",
    y = "Step count"
  ) + 
  geom_vline(
    xintercept = max_meanSteps_interval,
    col = "orange",
    lwd = 1.25
  ) + 
  geom_text(
    aes(
      label = ifelse(
        interval == max_meanSteps_interval,
        paste("Interval",as.character(interval)),
        ''
        )
      ),
    hjust=-0.1,
    vjust=0
  ) + 
  geom_text(
    aes(
      label = ifelse(
        interval == max_meanSteps_interval,
        paste(as.character(round(steps_by_interval)),"steps"),
        ''
        )
      ),
    hjust=-0.1,
    vjust=2
  )
```

### The 5-minute interval that, on average, contains the maximum number of steps
The previous plot depicts the the interval at which the most number of steps are
tracked on an average ie. at `r max_meanSteps_interval`, the user takes 
`r round(data_consdt$steps_by_interval[idx_max_meanSteps])` steps.

### Code to describe and show a strategy for imputing missing data
To impute missing values we'll use the mean step count data we've calculated 
previously and insert the mean value in the records with missing value with the 
corresponding mean value
```{r}
data_imputed = data
idx = which(is.na(data_imputed$steps))
for(i in idx){
  data_imputed$steps[i] = 
    round(
        data_consdt$steps_by_interval[which(
          data_imputed$interval[i] == data_consdt$interval
          )
        ]
    )
}
```

Imputed data
```{r}
head(data_imputed)
```

There are 0 missing values in the imputed dataset 'data_imputed'
```{r}
sum(is.na(data_imputed))
```

### Histogram of the total number of steps taken each day after missing values are imputed
Summarising the data
```{r}
data_grpd_imp = data_imputed %>% 
  group_by(date) %>%
  summarise(steps_per_day = sum(steps))
data_grpd_imp
```

Visualization
```{r}
ggplot(
  data_grpd_imp,
  aes(
    x = steps_per_day,
  ),
) + 
  geom_histogram(bins = 12, fill = "skyblue", col = "white") +
  stat_bin(bins = 12,geom="text", aes(label=..count..), col ="white", vjust = 1.3) +
  theme_bw() + guides(fill = F) + 
  labs(
    title = "Histogram of the total number of steps taken each day in the imputed data",
    x = "Steps taken each day",
    y = "Frequency")
```
### Comparing the average number of steps taken per interval across weekdays and weekends

Panel plot comparing the average number of steps taken per 5-minute interval 
across weekdays and weekends

Adding a column which indicates whether the date represents weekend or not
```{r}
library(lubridate)
data_imputed$weekend = factor(ifelse(wday(data_imputed$date) %in% c(2:6),0,1))
```

Considating the data by grouping by weekend first then interval
```{r}
data_consdt_imp = data_imputed %>% 
  group_by(weekend,interval) %>%
  summarise(steps_by_interval = mean(steps))
data_consdt_imp
```

Visualizing the weekday vs weekend step-count through the day
```{r}
levels(data_consdt_imp$weekend) = c("Weekday","Weekend")
ggplot(
  data_consdt_imp,
  aes(
    x = interval,
    y = steps_by_interval,
  )
) + 
  geom_line(col = 'skyblue', lwd = 1.25) + 
  theme_bw() + guides(fill = F) +
  labs(
    title = "Comparison between weekday and weekend daily step count mean",
    x = "Time interval of day",
    y = "Step count"
  ) + 
  facet_wrap(.~weekend, ncol = 1)
  
```


