---
title: "Storm Effects on Communities, Analysis"
author: "Anandu R"
date: "8/6/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Storms and other severe weather events can cause both public health and economic
problems for communities and municipalities. Many severe events can result in
fatalities, injuries, and property damage, and preventing such outcomes to the 
extent possible is a key concern.

This project involves exploring the U.S. National Oceanic and Atmospheric
Administration's (NOAA) storm database. This database tracks characteristics of 
major storms and weather events in the United States, including when and where 
they occur, as well as estimates of any fatalities, injuries, and property
damage.

# 1. Data Processing 
There is also some documentation of the database available. Details on how some 
of the variables are constructed/defined is available on this website by
National Weather Service : [Storm Data Documentation]("https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf")

## 1.1 Getting the data 
```{r message=FALSE, warning=FALSE, results='hide'}
fileUrl = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
if(!file.exists("./data/data.csv.bz2")){
  download.file(fileUrl,"./data/data.csv.bz2")
}
```  
## 1.2 Reading the data
```{r message=FALSE, warning=FALSE, results=F}
suppressMessages(library(dplyr))
data_raw <- read.csv("./data/data.csv.bz2", sep =",", header = T)
```

## 1.3 Preliminary analysis of data

```{r}
head(data_raw)
```

### 1.3.1 Reading column names 
```{r}
names(data_raw)
```


## 1.4 Data Cleaning   
### 1.4.1 Removing unnecessary variables/Subsetting the data  
Since the END_DATE and END_TIME fields are same as the BGN_DATA and BGN_TIME, we 
also remove those columns from the data.

Furthermore, since the COUNTY_END field has only the value 0 and would serve no
purpose to the analysis, it too is removed

The "REFNUM" and "REMARKS" fields don't serve any purpose to our analysis
```{r}
data_clean = select(data_raw, 
                    STATE,
                    COUNTY,
                    BGN_DATE,
                    BGN_TIME,
                    EVTYPE, 
                    FATALITIES,
                    INJURIES,
                    PROPDMG, 
                    PROPDMGEXP,
                    CROPDMG,
                    CROPDMGEXP)
```


### 1.4.2 Missing data treatment
#### 1.4.2.1 Checking distribution of Missing data and NAs in the dataset 
```{r}
as.numeric(colMeans(is.na(data_clean)))
as.numeric(colMeans(data_clean==""))
```

Columns 9, and 11 represent the "PROPDMGEXP", "CROPDMGEXP" fields which are
required for the analysis therefore we will keep them.

Therefore all in all, there arent any records to be removed or are their any 
columns that can be removed.

NOTE: During analysis there may still be some fields with no value aka missing
values in certain columns, but their percentages are in range 10-50% so the next
suitable step would be to  impute the values in the dataset, but since it is the
weather data, imputing values would only create noise in the data(?)

Looking at cleaned data 
```{r}
head(data_clean)
```

### 1.4.3 Fixing the datatypes and datafields  
#### 1.4.3.1 Creating a datatime field
```{r, cache= T}
data_clean$BGN_DATE = 
  as.POSIXct(data_clean$BGN_DATE, format = "%m/%d/%Y %H:%M:%S")

data_clean$BGN_TIME = 
  format(strptime(data_clean$BGN_TIME,"%H%M"),'%H:%M')

data_clean$BGN_DATETIME = 
  as.POSIXct(paste(data_clean$BGN_DATE,
                   data_clean$BGN_TIME
                   ), format="%Y-%m-%d %H:%M")
data_clean = 
  select(data_clean,
         STATE, COUNTY,
         BGN_DATETIME, 
         EVTYPE, FATALITIES,
         INJURIES,
         PROPDMG, 
         PROPDMGEXP,
         CROPDMG, 
         CROPDMGEXP)
```

#### 1.4.3.2 Imputing proper values in the "PROPDMGEXP", "CROPDMGEXP" fields
Current values in "CROPDMGEXP"
```{r}
unique(data_clean$CROPDMGEXP)
```
Current values in "PROPDMGEXP"
```{r}
unique(data_clean$PROPDMGEXP)
```
Correct representations:  
- "\"\"" = 10^0,  
- "-" = 10^0,   
- "?" = 10^0,  
- "+" = 10^0,  
- "0" = 10^0,  
- "1" = 10^1,  
- "2" = 10^2,  
- "3" = 10^3,  
- "4" = 10^4,  
- "5" = 10^5,  
- "6" = 10^6,  
- "7" = 10^7,  
- "8" = 10^8,  
- "9" = 10^9,  
- "H" = 10^2,  
- "K" = 10^3,  
- "M" = 10^6,  
- "B" = 10^9  

Imputing the correct values 
```{r}
data_clean = transform(data_clean,
                       PROPDMGEXP = toupper(PROPDMGEXP),
                       CROPDMGEXP = toupper(CROPDMGEXP))
DmgExP =  c("\"\"" = 10^0,
            "-" = 10^0, 
            "+" = 10^0,
            "?" = 10^0,
            "0" = 10^0,
            "1" = 10^1,
            "2" = 10^2,
            "3" = 10^3,
            "4" = 10^4,
            "5" = 10^5,
            "6" = 10^6,
            "7" = 10^7,
            "8" = 10^8,
            "9" = 10^9,
            "H" = 10^2,
            "K" = 10^3,
            "M" = 10^6,
            "B" = 10^9)
data_clean = transform(
  data_clean, 
  PROPDMGEXP = as.numeric(DmgExP[as.character(data_clean[,"PROPDMGEXP"])]),
  CROPDMGEXP = as.numeric(DmgExP[as.character(data_clean[,"CROPDMGEXP"])])
)
data_clean = transform(
  data_clean,
  PROPDMGEXP = ifelse(is.na(PROPDMGEXP),10^0,PROPDMGEXP),
  CROPDMGEXP = ifelse(is.na(CROPDMGEXP),10^0,CROPDMGEXP)
)
```


### 1.4.3.3 Subsetting the data, removing EVTYPEs that have 0 impact of any sort

```{r}
data_clean = subset(data_clean, 
                    EVTYPE != "?" &
                      (INJURIES > 0 | 
                         FATALITIES > 0 |
                         PROPDMG > 0 | 
                         CROPDMG > 0)
                    )
```

Looking at cleaned data 
```{r}
head(data_clean)
```

### 1.4.4 Standardising data in the '"EVTYPE" field

```{r}
unique(data_clean$EVTYPE)[1:10]
```

#### 1.4.4.1 Labeling event types correctly 
```{r echo=F, message=FALSE, warning=FALSE}
## WIND 1
data_clean[data_clean$EVTYPE=="NON TSTM WIND","EVTYPE"] = "WIND"
data_clean[data_clean$EVTYPE=="NON-TSTM WIND","EVTYPE"] = "WIND"

## THUNDERSTORM
data_clean[grepl("thunderstorm",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "THUNDERSTORM"
data_clean[grepl("thunderestorm",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "THUNDERSTORM"
data_clean[grepl("thundeerstorm",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "THUNDERSTORM"
data_clean[grepl("thunerstorm",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "THUNDERSTORM"
data_clean[grepl("THUNDERTORM WINDS",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "THUNDERSTORM"
data_clean[grepl("TUNDERSTORM WIND",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "THUNDERSTORM"
data_clean[grepl("THUDERSTORM WINDS", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "THUNDERSTORM"
data_clean[grepl("THUNDERSTROM WIND", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "THUNDERSTORM"
data_clean[grepl("tstm", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "THUNDERSTORM"

## WATERSPOUT + TORNADO 
data_clean[grepl("WATERSPOUT[\" \",/,-,+]TORNADO",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "WATERSPOUT+TORNADO"
data_clean[grepl("WATERSPOUT-TORNADO", 
                 data_clean$EVTYPE),"EVTYPE"] = "WATERSPOUT+TORNADO"
data_clean[grepl("WATERSPOUT/ TORNADO", 
                 data_clean$EVTYPE),"EVTYPE"] = "WATERSPOUT+TORNADO"
data_clean[grepl("WATERSPOUT", 
                 data_clean$EVTYPE),"EVTYPE"] = "WATERSPOUT+TORNADO"

## TORNADO 
data_clean[grepl("^TORNADO", 
                 data_clean$EVTYPE),"EVTYPE"] = "TORNADO"
data_clean[grepl("TORNDAO",
                 data_clean$EVTYPE),"EVTYPE"] = "TORNADO"
data_clean[grepl("FUNNEL CLOUD",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "TORNADO"
data_clean[grepl("COLD AIR TORNADO",
                 data_clean$EVTYPE),"EVTYPE"] = "TORNADO"

## LANDSLIDE 
data_clean[grepl("landslide", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "LANDSLIDE"

## FLASH FLOOD
data_clean[grepl("FLASH FLOOD", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "FLASH FLOOD"
data_clean[grepl("flash*FLOOD", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "FLASH FLOOD"
data_clean[grepl("FLOOD[/,\" \"]flash",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "FLASH FLOOD"

## COASTAL FLOOD
data_clean[grepl("COASTAL FLOOD",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "COASTAL FLOOD+EROSION"
data_clean[grepl("COASTAL  FLOODING/EROSION",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "COASTAL FLOOD+EROSION"
data_clean[grepl("Erosion/Cstl Flood",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "COASTAL FLOOD+EROSION"
data_clean[grepl("Erosion",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "COASTAL FLOOD+EROSION"

## FLOODS 
data_clean[grepl("FLOODING",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "FLOOD"
data_clean[grepl("FLOODS", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "FLOOD"
data_clean[grepl("RAPIDLY RISING WATER",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "FLOOD"

## OTHER FLOODS
data_clean[grepl("URBAN",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "URBAN FLOOD"
data_clean[grepl("RIVER", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "RIVER FLOOD"
data_clean[grepl("FLOOD/RAIN/WINDS",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "FLOOD"
data_clean[grepl("HEAVY RAIN AND FLOOD",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "FLOOD"
data_clean[grepl("HEAVY SNOW/HIGH WINDS & FLOOD",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "FLOOD"
data_clean[grepl("FLOOD & HEAVY RAIN",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "FLOOD"
data_clean[grepl("Ice jam flood \\(minor",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "FLOOD"
data_clean[grepl("LAKE FLOOD",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "RURAL FLOOD"
data_clean[grepl("LAKESHORE FLOOD", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "RURAL FLOOD"
data_clean[grepl("MAJOR FLOOD",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "URBAN FLOOD"
data_clean[grepl("RIVER FLOOD",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "RURAL FLOOD"
data_clean[grepl("SMALL STREAM FLOOD", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "RURAL FLOOD"

## TIDE 
data_clean[grepl("TIDE", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "TIDE"

## AVALANCHE
data_clean[grepl("Avalanche",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "AVALANCHE"
data_clean[grepl("Avalance", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "AVALANCHE"

## ICE SNOW BLIZZARD 
data_clean[grepl("FROST",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "FROST+SNOW"
data_clean[grepl("FREEZE", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "FROST+SNOW"
data_clean[grepl("COLD", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "FROST+SNOW"
data_clean[grepl("snow", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "FROST+SNOW"
data_clean[grepl("chill", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "FROST+SNOW"
data_clean[grepl("low temp",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "FROST+SNOW"
data_clean[grepl("winter",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "BLIZZARD"
data_clean[grepl("blizzard",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "BLIZZARD"
data_clean[grepl("ice storm", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "BLIZZARD"
data_clean[grepl("ice",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "FROST+SNOW"
data_clean[grepl("wintr",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "FROST+SNOW"
data_clean[grepl("freez",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "FROST+SNOW"
data_clean[grepl("sleet", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "FROST+SNOW"
data_clean[grepl("^glaze",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "FROST+SNOW"

## HEAT 
data_clean[grepl("HEAT",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "HEAT+DROUGHT"
data_clean[grepl("WARM", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "HEAT+DROUGHT"
data_clean[grepl("DROUGHT", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "HEAT+DROUGHT"

## DUST 
data_clean[grepl("DUST", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "DUST"

## WILDFIRE 
data_clean[grepl("FIRE", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "WILDFIRE"

## HURRICANE 
data_clean[grepl("hurricane", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "HURRICANE"

## HAIL
data_clean[grepl("HAIL", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "HAIL"

## SURF 
data_clean[grepl("surf", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "SURF"

## WIND 2
data_clean[grepl("wind",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "WIND"
data_clean[grepl("burst", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "WIND"

## MUDSLIDES
data_clean[grepl("mud", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "MUD+LAND SLIDES"
data_clean[grepl("land", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "MUD+LAND SLIDES"
data_clean[grepl("rock", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "MUD+LAND SLIDES"

## RAINFALL 
data_clean[grepl("rain", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "RAINFALL"
data_clean[grepl("precip",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "RAINFALL"
data_clean[grepl("show",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "RAINFALL"

## LIGHTNING 
data_clean[grepl("light", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "LIGHTNING"
data_clean[grepl("ligntning", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "LIGHTNING"

## TROPICAL STORM
data_clean[grepl("tropical",
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "TROPICAL CYCLONE"

## SURGE
data_clean[grepl("surge", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "STORM SURGE"

## SLEET
data_clean[grepl("sleet",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "SLEET"

## RIP CURRENT
data_clean[grepl("rip", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "RIP CURRENT"

## GUST
data_clean[grepl("gust", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "WIND"

## HYPOTHERMIA
data_clean[grepl("hypo", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "HYPOTHERMIA"

## HYPERTHERMIA
data_clean[grepl("hyper", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "HYPERTHERMIA"

## SWELLS + SEAS + MIX + HIGH WATER + WAVES 
## WET
data_clean[grepl("wet", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "WETNESS"

## TIDES
temp = unique(data_clean$EVTYPE)[c(20,27,39,40,42,43,44,54,57,58)]
data_clean[data_clean$EVTYPE %in% temp,"EVTYPE"] = "TIDE/ROUGH SEAS"

## SEICHE
data_clean[grepl("seiche", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "TIDE/ROUGH SEAS"

## COASTAL STORM 
data_clean[grepl("coastal storm",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "COASTAL STORM"
data_clean[grepl("coastalstorm", 
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "COASTAL STORM"

## MARINE MISHAP/ACCIDENT
data_clean[grepl("marine", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "MARINE ACCIDENT"

## OTHER
data_clean[grepl("^other", 
                 data_clean$EVTYPE,
                 ignore.case = T),"EVTYPE"] = "OTHER"

## APACHE COUNTY to WIND since mentioned in REMARKS 
data_clean[grepl("APACHE COUNTY",
                 data_clean$EVTYPE, 
                 ignore.case = T),"EVTYPE"] = "WIND"

```


```{r}
unique(data_clean$EVTYPE)
```


## 2. Exploratory Analysis 

### 2.1 Creating new fields CROPDMGPRICE and PROPDMGPRICE  
```{r}
data_clean = transform(data_clean, 
                       CROPDMGPRICE = CROPDMG*CROPDMGEXP,
                       PROPDMGPRICE = PROPDMG*PROPDMGEXP)
```


### 2.2 Aggregating the data based on event type
```{r, results=T}
library(dplyr)
suppressMessages(
  {
  data_aggr = data_clean %>%
    group_by(EVTYPE) %>%
    summarise(
      FATALITIES = sum(FATALITIES, na.rm = T),
      INJURIES = sum(INJURIES,na.rm = T),
      CROPDMGPRICE = sum(CROPDMGPRICE, na.rm = T),
      PROPDMGPRICE = sum(PROPDMGPRICE, na.rm = T)
    )
  }
)
head(data_aggr[order(-data_aggr[,"FATALITIES"],
                     -data_aggr[,"INJURIES"],
                     -data_aggr[,"CROPDMGPRICE"],
                     -data_aggr[,"PROPDMGPRICE"]),])
```
```{r}
nrow(data_aggr)
```  
There are 46 rows of data available on various events, which we'll use to create 
various plots to show which types of events across the United States are most 
harmful with respect to population health and have the greatest economic
consequences.

### 2.3 Analyis to find events most harmful with respect to population health

Looking at data in relevant columns "FATALITIES" and "INJURIES" sorted by 
descending order of the field values  
```{r}
head(data_aggr[order(-data_aggr[,"FATALITIES"], -data_aggr[,"INJURIES"]),
               c("EVTYPE","FATALITIES","INJURIES")])
```

### 2.4 Analyis to find events most harmful with respect to economic damage

Looking at data in relevant columns "CROPDMGPRICE" and "PROPDMGPRICE" sorted by 
descending order of the field values
```{r}
head(data_aggr[order(-data_aggr[,"CROPDMGPRICE"], -data_aggr[,"PROPDMGPRICE"]),
               c("EVTYPE","CROPDMGPRICE","PROPDMGPRICE")])
```

## 3. RESULTS 
### 3.1 Visualization   
#### 3.1.1 Population health  
```{r}
ggplot(
  data_aggr,
  aes(
    
  )
)
```


#### 3.1.2 Economic damage  



## Removing data file after analysis
```{r}
unlink("./data/data.csv.bz2",recursive = T)
#unlink("./analysis_cache", recursive = T)
```

