---
title: "Face Data Analysis"
author: "Anandu R"
date: "6/15/2020"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Face Data Analysis
Loading the data 
```{r}
load("face.rda")
image(t(faceData)[,nrow(faceData):1])
```
# Running SVD

```{r}
svdFace = svd(scale(faceData))
plot(svdFace$d^2/sum(svdFace$d^2),pch = 19,xlab = "Singular vector", ylab = "Variance Explained")
```
We observe that almost 40% of the variation is explained by the first singular vector, and almost all of the variation in the data is explained by the first 5 singular vectors 

We now use matrix multiplication to create an approximation of the face data with fewer components, preferably by using just the first 5 singular vectors 

```{r}
approx1 = svdFace$u[,1] %*% t(svdFace$v[,1]) * svdFace$d[1]
approx3 = svdFace$u[,1:3] %*% diag(svdFace$d[1:3]) %*% t(svdFace$v[,1:3])  
approx5 = svdFace$u[,1:5] %*% diag(svdFace$d[1:5]) %*% t(svdFace$v[,1:5])  
approx9 = svdFace$u[,1:9] %*% diag(svdFace$d[1:9]) %*% t(svdFace$v[,1:9])  
```

The above variables are approximation of the original face data using number of components as specified by the suffixed number 

Visualising and comparing the approximation with original data
```{r}
par(mfrow = c(1,3))
image(t(approx1)[, nrow(approx1):1], main = "1 vector")
image(t(approx5)[, nrow(approx5):1], main = "5 vectors")
image(t(faceData)[, nrow(faceData):1], main = "Original data")
```

Comparing the gain/loss in details for various number of singular vectors used
```{r}
par(mfrow = c(1,3))
image(t(approx3)[, nrow(approx3):1], main = "3 vectors")
image(t(approx5)[, nrow(approx5):1], main = "5 vectors")
image(t(approx9)[, nrow(approx9):1], main = "9 vectors")
```




