---
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height=5, fig.width=5, fig.align = "centred")
```

# Motor Trend: Automatic vs manual transmission comparative study   
You work for Motor Trend, a magazine about the automobile industry. Looking at a
data set of a collection of cars, they are interested in exploring the 
relationship between a set of variables and miles per gallon (MPG) (outcome). 
They are particularly interested in the following two questions:  
- “Is an automatic or manual transmission better for MPG”  
- “Quantify the MPG difference between automatic and manual transmissions”  

Data | Description
--- | ---
mpg| Miles per US gallon                        
cyl| Number of cylinders                      
disp| Displacement (cubic inches)                    
hp|Gross horsepower                         
drat| Rear axle ratio                          
wt| Weight (lb / 1000)                         
qsec| 1 / 4 mile time                            
vs| V/S                                      
am| Transmission (0 = automatic, 1 = manual) 
gear| Number of forward gears                  
carb| Number of carburetors


## 1.Loading prerequisites  
#### 1.1. Libraries  
```{r}
suppressMessages(
    {
        if(!require(manipulate)){
            install.packages("manipulate")
        }
        if(!require(GGally)){
            install.packages("GGally")
        }
        if(!require(lmtest)){
            install.packages("lmtest")
        }
        if(!require(dplyr)){
            install.packages("dplyr")
        }
        if(!require(ggplot2)){
            install.packages("ggplot2")
        }
        library(GGally)
        library(manipulate)
        library(lmtest)
        library(dplyr)
        library(ggplot2)
    }
)
```

#### 1.1. data  
```{r}
data(mtcars)
head(mtcars)
```  

## 2. Exploratory Analysis    
**Variables**  
```{r}
names(mtcars)
```
**Understanding variable types**  
```{r}  
apply(mtcars,2,class)
```
**Fixing the types**  
```{r}
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am <- factor(mtcars$am,labels=c('Automatic','Manual'))
```


**Summarizing each variables**  
```{r}
summary(mtcars)
``` 

## 3. Regression modelling   
Fitting a model with all the variables   
```{r}
mdl_all = glm(mpg~., family = "gaussian", data = mtcars)
summary(mdl_all)
```
The above model tells us that the average **mpg** is `r round(summary(mdl_all)$coef[1,1],2)`  

#### 3.1 Model selection  
- We would want to select a model that has larger adjusted and predicted R-squared
values.
- In regression, p-values less than the significance level indicate that the 
term is statistically significant. We reduce the model by repeatedly removing 
parameters corresponding to coefficients that do not have significant effect on 
the model performance.  
- Upon arriving at the manually identified "best" model, we will compare the 
model performance against an automated model selection procedure using step 
regression function. step()  
- We will analyse whether there is significant effect on mpg when considering 
the transmission model using the t.test() function.

The wt, carb and disp parameters has least significant effects on the model 
performance, Hence we remove them to analyse the variation in the p-value of the
other variables. If there is large improvements then it'd mean that there is no
correlation between them.
```{r}
mdl_test_1 = glm(mpg~.-wt-disp-carb, family = "gaussian", data = mtcars)
summary(mdl_test_1)
```
Similary we remove qsec, gear, drat and vs
```{r}
mdl_test_2 = glm(mpg~.-wt-disp-carb-qsec-gear-drat-vs, family = "gaussian", data = mtcars)
summary(mdl_test_2)
```  

Automated best model selection using step regression  
```{r results=F, warning=F, message=F}
mdl_best = step(mdl_all, direction = "backward")
```
```{r}
summary(mdl_best)
```

The two models seem to be very similar with exception of just the wt parameter
being omitted from the manually subsetted model.

#### 3.2 Analyzing the models  

Analysing the residual plots manually fitted vs step regressed.
```{r fig.height=4, fig.width=7}
par(mfrow=c(1,2))
plot(mdl_test_2, which =1)
plot(mdl_best, which =1)
```  
It is observed that the manually fitted model has lower standard error.

**Asserting the normality of the residuals of each models**
```{r fig.height=4, fig.width=7}
par(mfrow=c(1,2))
plot(mdl_test_2, which =2)
plot(mdl_best, which =2)
```


#### 3.3 Using T-test to identify significance of transmission type on mpg
```{r}
t.test(mpg ~ am, data = mtcars)
```
The above t.test with p-statistic <0.05 shows that there is significant effect 
of type of transmission on the mpg of a motor vehicle.  

Visualizing the difference
```{r}
ggplot(
  mtcars,
  aes(
    x=am,
    y=mpg,
    fill = am
  )
) + geom_boxplot() +
  labs(x = "Transmission type", y = "miles per gallon") + 
  ggtitle("Effect of transmission type on mpg of motor vehicles") +
  theme_bw() 
```  

## 4. Conclusion
Based on my analysis I was able to identify that there is significant effect of 
transmission type of the motor vehicle on its miles per gallon metric - that the 
manual transmission performes better than automatic type.  

The rate of change of the conditional mean mpg with respect to am is about 4.1 
conidering the manually subsetted model and 1.8 with the best subset selection 
using the step function.  

Using the T-test we were able to infer that there is indeed a significance in 
transmission type on mpg visualized using boxplot shown above




