---
title: "ToothGrow Analysis"
author: "Anandu R"
date: "9/20/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=5, fig.height=3, fig.align = "center")
library(ggplot2)
library(dplyr)
```


## ToothGrowth data analysis
Loading the data
```{r}
data("ToothGrowth")
```

Summarizing the data
```{r}
ToothGrowth$dose = factor(ToothGrowth$dose)
str(ToothGrowth)
```


## Exploratory analysis using boxplots
### Mean change in tooth length vs supplement delivery mode
```{r, echo = F}
ggplot(
  ToothGrowth,
  aes(
    x = supp,
    y = len,
    fill = supp
  )
) + geom_boxplot() + 
  ggtitle("Tooth length vs Supplement") + 
  theme_bw()
```   

There is considerable difference in the mean difference in tooth growth observed
in the patients for the different delivery methods of supplement provided

### Difference in the tooth length for various dosages
```{r, echo = F}
ggplot(
  ToothGrowth,
  aes(
    x = dose,
    y = len,
    fill = dose
  )
) + geom_boxplot() +
  ggtitle("Tooth length vs Dosage") +
  theme_bw()
```    

As dosage is increased the average tooth length increases at each step increase 
in dosage.

### Mean change in tooth length analyzed across each supplement and dosage provided
```{r, echo = F}
ggplot(
  ToothGrowth,
  aes(
    x = dose,
    y = len,
    group = dose,
    fill = supp
  )
) + geom_boxplot() +
  facet_wrap(supp~.) + 
  ggtitle("Supplement~Dosage effect on Toothlength") +
  theme_bw()
```  

We observe that there is marked difference mean distribution of tooth length for
patients that consumed 0.5 or 1 dosage of supplement OJ and VC, but the 
difference is almost non-existent for the 2mg dosage.  

## Hypothesis testing 
### Testing by supplement delivery mode
Consider the supplement OJ being better performing than supplement VS
```{r}
## Subsetting the data
x = ToothGrowth$len
group = ToothGrowth$supp
## Perfomring t-test
t.test(x~group, paired = F)
```  
The p value is greater than <img src="https://render.githubusercontent.com/render/math?math=$\alpha=0.05$"> hence we reject the null hypothesis

Calculating the power with which we can reject the null hypothesis
```{r}
n = length(x)
mu0 = mean(x[group=="OJ"])
mua = mean(x[group=="VC"])
sigma = sd(x[group=="VC"])
delta = (mua - mu0)/sigma
power.t.test(n = n, sd = sigma, delta = delta, type = "one.sample", alt = "one.sided")$power
```  
Our power is pretty low but regardless, the null hypothesis has been rejected
but we can also bring out due to the low power that there isn't much significant
difference between the supplement delivery modes.

```{r, echo = F}
df = data.frame(lab=group,means=x)
ggplot(
  df,
  aes(
    x=means,
    fill = lab
  )
) + geom_density(alpha = 0.5) + 
  theme_bw() + 
  ggtitle("Distribution of lengths for delivery modes")
```  
As we can see from the visualization that the two distributions are highly 
overlapping.

### Testing by dosage
Another thing we noticed from the supplement~dosage visualization was that there 
is considerable difference in output for the various dosage levels 0.5, 1 and 2.  
Consider the null hypothesis being dosage level has no effect on the tooth 
length.

Let us now check to see if we get a significant p-value for each pairing of the 
dosage.
```{r}
x = subset(ToothGrowth ,dose %in% c(0.5,1))$len
group = subset(ToothGrowth, dose %in% c(0.5,1))$dose
t.test(x~group, paired = F)
```  
```{r, echo = F}
df = data.frame(lab=group,means=x)
ggplot(
  df,
  aes(
    x=means,
    fill = lab
  )
) + geom_density(alpha = 0.5) + 
  theme_bw() +
  ggtitle("Distribution of lengths in groups 0.5 and 1")
```  


```{r}
x = subset(ToothGrowth ,dose %in% c(1,2))$len
group = subset(ToothGrowth, dose %in% c(1,2))$dose
t.test(x~group, paired = F)
```  
```{r, echo = F}
df = data.frame(lab=group,means=x)
ggplot(
  df,
  aes(
    x=means,
    fill = lab
  )
) + geom_density(alpha = 0.5) + 
  theme_bw() +
  ggtitle("Distribution of lengths in groups 1 and 2")
```  


```{r}
x = subset(ToothGrowth ,dose %in% c(0.5,2))$len
group = subset(ToothGrowth, dose %in% c(0.5,2))$dose
t.test(x~group, paired = F)
```  
```{r, echo = F}
df = data.frame(lab=group,means=x)
ggplot(
  df,
  aes(
    x=means,
    fill = lab
  )
) + geom_density(alpha = 0.5) + 
  theme_bw() + 
  ggtitle("Distribution of lengths in groups 0.5 and 2")
```  


In each case the p-value is very low hence we can conclude that the hypothesis 
can be easily rejected. 

## Conclusions
There appears to be statistical significance on tooth growth by varying the 
dosage levels, as dosage increases the mean tooth length increases.  
And, There seems to be negligible role played by the delivery methods - VC, OJ 
although it is to be noted that at dosage levels 0.5 and 1 OJ has higher 
overall outcome in toothgrowth, whereas negligible for dosage of level 2



















