---
title: "ggplot"
author: "Anandu R"
date: "6/15/2020"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dimension Reduction

Principal component analysis (PCA) and singular value decomposition (SVD) are commonly used dimensionality reduction approaches in exploratory data analysis (EDA) and Machine Learning. They are both classical linear dimensionality reduction methods that attempt to find linear combinations of features in the original high dimensional data matrix to construct meaningful representation of the dataset.

# Illustrating clusters 

```{r}
set.seed(96)
x = rnorm(12, mean = rep(1:3, each = 4), sd = 0.2)
y = rnorm(12, mean = rep(c(1,2,1),each = 4), sd = 0.2)
plot(x, y, col = "orange", pch = 19,cex = 1.5)
text(x+0.05, y+0.05, labels = as.character(1:12))
```

# Missing values

It is imporatant to deal with missing values when working with SVD, since error occurs if our data contains any sort of missing value. 

One way to deal with errors is to use built-in R function 'impute()'

Consider the dataset 
```{r}
x = rnorm(12, mean = rep(1:3, each = 4), sd = 0.2)
y = rnorm(12, mean = rep(c(1,2,1),each = 4), sd = 0.2)
z = x[sample(1:12)]
w = y[sample(1:12)]
data = data.frame(x,y,z,w)
dataAsMatrix = as.matrix(data)[sample(1:12),]
image(t(dataAsMatrix)[,nrow(dataAsMatrix):1],yaxt = "n")
```

Creating new object 'dataMatrixOrdered' - data ordered according to the hierarchical clusters.
```{r}
library(dplyr)
hh <- dist(dataAsMatrix) %>% hclust
dataMatrixOrdered <- dataAsMatrix[hh$order, ]
image(t(dataMatrixOrdered)[,nrow(dataMatrixOrdered):1],yaxt = "n")
```


Loading the necessary impute package, to impute the missing values to perform dimension reduction

```{r}
if(!require(impute)){
  if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
  BiocManager::install("impute")
}
```

# Imputing 
This function uses K-nearest neighbours and imputes missing value to a row with the value of the nearest neighbour
```{r}
set.seed(19)
library(impute)
dataMatrix2 = dataMatrixOrdered
# Randomly insert some missing data to test the impute function performance
dataMatrix2[sample(1:48,size=6,replace = F)] <- NA
dataMatrix2 = impute.knn(dataMatrix2)$data
```

Comparing the SVD analysis between original data and imputed data
```{r}
svdOg = svd(scale(dataMatrixOrdered))
svdImp = svd(scale(dataMatrix2))
par(mfrow = c(1,2))
plot(svdOg$v[,1],pch=19)
plot(svdImp$v[,1],pch=19)
```


