---
title: "ggplot"
author: "Anandu R"
date: "6/15/2020"
output: pdf_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# K-Means Clustering 

Illustrating clusters 

```{r}
set.seed(96)
x = rnorm(12, mean = rep(1:3, each = 4), sd = 0.2)
y = rnorm(12, mean = rep(c(1,2,1),each = 4), sd = 0.2)
plot(x, y, col = "orange", pch = 19,cex = 1.5)
text(x+0.05, y+0.05, labels = as.character(1:12))
```

### Using the kmeans() 

This function is used to find various clusters in the dataset, it returns an object with various elements
the element "cluster" gives a list of integers denoting to which cluster each record in the datset belongs to

```{r}
data = data.frame(x,y)
kmeansobj = kmeans(data, centers = 3)
names(kmeansobj)
```

```{r}
kmeansobj$cluster
```

The coordinates of centroids of each cluster is stored in the "centers" element
```{r}
kmeansobj$centers
```


### Plotting the data using the identified clusters 

```{r}
plot(x, y, col = (kmeansobj$cluster+1), pch = 19, cex = 1.5)
points(kmeansobj$centers, pch = 3, col = 2:4, cex = 1.5, lwd = 1.5)
```

## Generating heatmaps 

```{r}
dataAsMatrix = as.matrix(data)[sample(1:12),]
kmeansobj_new = kmeans(dataAsMatrix,centers = 3)
image(t(dataAsMatrix)[,nrow(dataAsMatrix):1],yaxt = "n")
```

To reorder the data such that the records of same clusters are together

```{r}
image(t(dataAsMatrix)[,order(kmeansobj_new$cluster)],yaxt = "n")
```

