---
title: "Air Quality PM2.5 Case Study"
author: "Anandu R"
date: "7/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Air Quality PM2.5 Case Study

## Loading the data

```{r}
fileUrl2012 = 
  "https://aqs.epa.gov/aqsweb/airdata/annual_conc_by_monitor_2012.zip"
fileUrl1999 = 
  "https://aqs.epa.gov/aqsweb/airdata/annual_conc_by_monitor_1999.zip"
if(!file.exists('./data_1999.csv')&&!file.exists('./data_2012.csv')){
  invisible(download.file(fileUrl1999,"./data_1999.zip"))
  invisible(download.file(fileUrl2012,"./data_2012.zip"))
  unzip("data_2012.zip", exdir = getwd())
  unzip("data_1999.zip", exdir = getwd())
}
unlink('./data_1999.zip')
unlink('./data_2012.zip')
tryCatch(
  invisible(
    file.rename(
      "annual_conc_by_monitor_1999.csv","data_1999.csv"
      )
  ),warning = function(w){
      if(file.exists('./data_1999.csv')){
        print("The file \"data_1999.csv\" already exists!")
      }else{
        print("Some error occured while trying to rename the file \"annual_conc_by_monitor_1999.csv\" to \"data_1999.csv\"!")
      }
    }
  )
tryCatch(
  invisible(
    file.rename(
      "annual_conc_by_monitor_2012.csv","data_2012.csv"
      )
  ),warning = function(w){
      if(file.exists('./data_1999.csv')){
        print("The file \"data_2012.csv\" already exists!")
      }else{
        print("Some error occured while trying to rename the file \"annual_conc_by_monitor_2012.csv\" to \"data_2012.csv\"!")
      }
    }
  )
data_1999 = read.csv("data_1999.csv")
data_2012 = read.csv("data_2012.csv")
```

## Initial analysis on the data

Identifying the parameter code for PM2.5 data
```{r}
invisible(library(dplyr))
par_code = grep("PM2.5",data_1999$Parameter.Name)[1]
par_code = data_1999$Parameter.Code[par_code]
```

Filtering dataset with only the required parameter list - PM2.5
```{r}
data_1999 = filter(data_1999, Parameter.Code == par_code)
data_2012 = filter(data_2012, Parameter.Code == par_code)
```

Size of the dataset
```{r}
sprintf("1999 data, Rows = %d, Columns = %d",dim(data_1999)[1],dim(data_1999)[2])
sprintf("2012 data, Rows = %d, Columns = %d",dim(data_2012)[1],dim(data_2012)[2])
```
Subsetting just the particulate matter measure column from each dataset
```{r}
pm_1999 = data_1999$X1st.Max.Value
pm_2012 = data_2012$X1st.Max.Value
```

Structure of variable
```{r}
summary(pm_1999)
summary(pm_2012)
```
0% of the data is missing 
```{r}
mean(is.na(pm_1999))
mean(is.na(pm_2012))
```

## Visualization
boxplot
```{r}
boxplot(log10(pm_1999),log10(pm_2012))
```
We observe that eventhough the average values have gone down, the spread of the
data has increased to the right extreme, i.e the data has become more right 
skewed in 2012, from being left skewed in 1999.

Extracting the dates in which the measurements were taken
```{r}
date_1999 = as.Date(data_1999$X1st.Max.DateTime)
date_2012 = as.Date(data_2012$X1st.Max.DateTime)
str(date_1999)
str(date_2012)
```
```{r}
par(mfrow = c(1,2))
hist(date_1999,"month")
hist(date_2012,"month")
```
To make analysis more simpler we pick a common monitor from 1999 and 2012, and 
compare the pm2.5 levels between the two instead of the levels of the entire 
country. This will also allow us to control for possible changes in the
monitoring locations between 1999 and 2012. As new monitors are added overtime 
and this can lead to inaccuracy in the analysis by taking into consideration the 
entire country.

Now we subset the data to look at the various monitors present in New York City
```{r}
m_NY_1999 = unique(subset(data_1999, State.Code==36, c(County.Code, Site.Num)))
m_NY_2012 = unique(subset(data_2012, State.Code==36, c(County.Code, Site.Num)))

## Creating a new variable that pastes the values of County.Code and Site.Num 
## inorder to find the intersecting County-Site combinations in both time periods
## within newyork city
m_NY_1999 = paste(m_NY_1999[,1],m_NY_1999[,2],sep = ".")
m_NY_2012 = paste(m_NY_2012[,1],m_NY_2012[,2],sep = ".")

## Finding the intersecting county-site combinations
common_m = intersect(m_NY_1999,m_NY_2012)
common_m
```
Thus we're able to find the common monitor we can use for out analysis 
Now we need to select the monitor that has the most number of observations, 
since more data equals better analysis and visualizations

Subsetting the data from NY to only have records of observations that are from
the common monitors 
```{r}
## Creating a new variable county.site, to subset 
m_NY_1999 = data_1999
m_NY_2012 = data_2012
m_NY_1999$County.Site = paste(m_NY_1999$County.Code,m_NY_1999$Site.Num,sep = ".")
m_NY_2012$County.Site = paste(m_NY_2012$County.Code,m_NY_2012$Site.Num,sep = ".")
m_NY_1999 = subset(m_NY_1999, County.Site %in% common_m)
m_NY_2012 = subset(m_NY_2012, County.Site %in% common_m)
```

Looking at the number of observations for each County.Site variable
```{r}
sapply(split(m_NY_1999, m_NY_1999$County.Site),nrow)
sapply(split(m_NY_2012, m_NY_2012$County.Site),nrow)

```
Picking the county.site combination 1.5 i.e county 1, and site 5 within that 
county.

```{r}
m_1999 = subset(data_1999, State.Code == 36 & County.Code == 1 & Site.Num == 5)
m_2012 = subset(data_1999, State.Code == 36 & County.Code == 1 & Site.Num == 5)
```





## Deleting files after completion
```{r}
unlink("data_1999.csv")
unlink("data_2012.csv")
```


