---
title: "EDA Human Activity Analysis using Smartphone Dataset"
author: "Anandu R"
date: "7/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Setting proper Working directory for the project 
```{r}
## Saving main directory path
setwd_future = getwd()

## Navigating to project directory
setwd(
  "Case Studies/Human_Activity_Recognition_Using_Smartphones_Data_Set/Cleaning"
  )

## Saving location of project directory
setwd_UCI_HAR = getwd()

```

# EDA Human Activity Analysis using Smartphone Dataset

The dataset has information on various sensors on mobilephone, the dataset has
been split into training(70%) and testing(30%), with activities manually 
labelled with the associated activity and subject under consideration in the 
training dataset.

We use this data to predict the activity of the user based on readings from the 
mobile sensors

# Loading the data 
```{r}
fileUrl = 
  "http://archive.ics.uci.edu/ml/machine-learning-databases/00240/UCI%20HAR%20Dataset.zip"

if (!file.exists('./UCI HAR Dataset')){
  download.file(fileUrl,'./UCI HAR Dataset.zip', mode = 'wb')
  unzip("UCI HAR Dataset.zip", exdir = getwd())
}
unlink("./UCI HAR Dataset.zip")
unlink("./__MACOSX",recursive = T)
```



# Deleting unecessary folder
```{r}
unlink("./UCI HAR Dataset",recursive = T)
```


