---
title: "ggplot"
author: "Anandu R"
date: "6/9/2020"
output: pdf_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic components of ggplot2 Plot 
- A **data frame**
- **aesthetic mappings**: How data are mapped to color, size.
- **geoms**: Geometric objects like points, lines, shapes.
- **facets**: For conditional plots 
- **stats**: statistical transformations like binning, quantiles, smoothing.
- **scales**: What scale an aesthetic map uses(example, male = red, female = blue)
- **coordinate system**

## Making basic plot 

Importing necessary libraries 

```{r}
invisible(library(ggplot2))
data = mtcars
```

Creating plot object 

```{r}
plot = ggplot(
  data, aes(
    x = wt,
    y = mpg
  )
)
plot
```

The graph page has been made, but no data has been plotted on the graph yet, we need to specify the data 

```{r}
plot = plot + geom_point()
plot
```

Adding a smoother 

```{r}
plot + geom_smooth()
```

The above smoother is by default method = "loess", using a linear model - lm smoother 

```{r}
plot + geom_smooth(method = "lm", se = F)
```

Adding facets 

```{r}
plot + facet_grid(.~cyl) + geom_smooth(method = "lm",se = F)
```

## Annotations 

- Labels: xlab(), ylab(), ggtitle)(
