---
title: "qplot"
author: "Anandu R"
date: "6/2/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Visualizing the mtcars dataset using ggplot's qplot function

Importing necessary libraries 

```{r}
invisible(library(ggplot2))
invisible(library(datasets))
```

Modifying dataset  before plotting

```{r}
str(mtcars)
```

Converting cyl variable to factor

```{r}
mtcars$cyl = factor(mtcars$cyl)
```


Plotting the mpg (Miles per gallon) against weight, subsetting by number of cylinders

```{r}
qplot(mpg,wt,data = mtcars,color = cyl)
```

## Histograms 

Plotting a histogram to analyse the mpg (miles per gallon)

```{r}
qplot(mpg,data= mtcars,fill = cyl,bins = 20)
```

## geom loess function

Adding a geom to the plot 

```{r}
qplot(mpg,wt,data = mtcars,geom = c("point","smooth"))
```

## Facets 

Creating facets, which are essentially panels in ggplot, where plots are put in panels based on a factor variable

```{r}
qplot(mpg,wt,data = mtcars, facets = .~cyl)
```


```{r}
qplot(mpg,data= mtcars,facets = cyl~.,bins = 20)

```
