---
title: "examples of PDFs"
author: "Anandu R"
date: "8/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Example 1

Consider the function to denote the probability of help calls addressed by a 
helpline be 
```
  f(x) = {2x for 0<x<1; 0 otherwise}
``` 
```{r, echo=FALSE}
x = c(-0.5,0,1,1,1.5)
y = c(0,0,2,0,0)
plot(x, y, type = "l", frame = F)
```   

To find probability that 75% or fewer of calls get addressed   

```{r, echo=FALSE}
plot(x, y, type = "l", frame = F)
lines(c(0,0.75,0.75),c(0,0,1.5), col = "skyblue")
```  

Find area of given portion 
```
  (1/2)*(0.75)*(1.5)
  = 0.5625
```

The same can be achieved by using the beta function for probability - pbeta()  
```{r, results=F}
pbeta(0.75,2,1)
```  
The quantile for the above population distribution  
Median from the distribution shows the datapoint below which 50% of the data is 
present and above it is the other 50%

```
  0.5 = F(X) = P(X<=x) = 50% of the area
  Integrating the function F(X)=2x, we get x^2
  => x = sqrt(2) = 0.707
  => The required datapoint
```  

The same can be found out using the qbeta() function which gives the quantile of
the beta distribution
```{r, results=F}
qbeta(0.5,2,1)
```

The inference is that during 50% of the day ~70% of the calls are addressed. 

  



